Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-07-26 Final_Manuscript.Rmd
added references

##Self-Objectification
Self-objectification is a multidimensional process that accounts for the cognitive mechanism that translates experiences of sexualization at the cultural level to psychological (e.g., anxiety, self-esteem, authenticity, motivational states) and behavioral (cognitive performance, body monitoring) features of mental health and well-being at the individual level [@calogero2011; @moradi2008]. @calogero2011 proposes that the construct of self-objectification can be conceptualized as a learned trait. Furthermore, it can also be elicited momentarily, through the media, for example, with sexualized images in movies and magazines, which can lead to a state of self-objectification [@calogero2011, @moradi2008]. Being objectified by another person and possessing trait-level self-objectification (TSO) may interact to influence experiences of feeling like a body, or state self-objectification (SSO) [CITE swimsuit sweater, @garcia2016objectification]. Does this habit self-objectification lead women to also objectify other women, possibly to compare oneself?

@@

##Interpersonal Objectification
SIMO model and literature review [@gervais2020social]. Studies have shown that within social encounters women are gazed at more than men [@briton1995], often times feel ‘looked at’ within interpersonal interactions [@argyle1969], and will more than likely internalize the objectifying gaze on physical self [@young1979]. Moreover, perhaps the most adverse effect of objectifying treatment is that it effectively socializes girls and women to treat themselves as objects to be looked at and evaluated, an effect termed self-objectification  [@Bartky; @berger1972; @fredricksonetal1998]. We don't know about women-on-women interpersonal objectification. What might be different?

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successful relationship, thus, the "college relationships condition may heighten self-objectification and the evaluation of other women in a sexualized way. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship and they experience less SSO after appearance comments within romantic relationships [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship building skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@quinnetal]. Both romantic relationships and relationships in general have been studies in the context of objectification, but neither of these literatures have investigated specifically same-gender relationships. LANEE: COMMENT IN ALL CAP LIKE THIS!

There is evidence that the expereince of state self-objectification in mixed-gender contexts (stranger and romantic) and within the self in any moment (trying on a swimsuit) has negative consequences, but what about in the context of interaction with other women? Is the self-objectification expreience in mixed-gender interactions the same as that experienced in same-gender interactions? The ample research demonstrating that the male gaze has a particularly detrimental effect would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my]. Women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], but when they do it, does it lead to self-ibjectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that ineraction?

##Interpersonal Objectification and Authenticity
There is some evidence that reduced authenticity is a consequence of self-objectification in the moment [@garcia2016objectification; again CITE PWQ paper I reviewed, Yoder editor]. This link has been justified by the lierature on stigmatized-stigmatizer interactions, viewing the experience of being objectified in an interaction as an identity threat situation. When a woman is objectified by a man, and then experiences self-objectification, this intergroup encounter might trigger identitiy threat, but is the same negative consequence present when a woman is objectified by another woman? woman-woman interpersonal objectification processes might diverge from this process. One the one hand, she may not be having the same negative consequences the cascade from situations that threated group-based identity [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in an interaction may be enough to reduce feelings of authenticity and social competnce, regardless of the gender of the objectifyer [@tolman2006looking; PWQ paper I reviewed, Yoder editor].

>>>>

##Self-Objectification
Self-objectification is a multidimensional process that accounts for the cognitive mechanism that translates experiences of sexualization at the cultural level to psychological (e.g., anxiety, self-esteem, authenticity, motivational states) and behavioral (cognitive performance, body monitoring) features of mental health and well-being at the individual level [@calogero2011; @moradi2008]. @calogero2011 proposes that the construct of self-objectification can be conceptualized as a learned trait. Furthermore, it can also be elicited momentarily, through the media, for example, with sexualized images in movies and magazines, which can lead to a state of self-objectification [@calogero2011, @moradi2008]. Being objectified by another person and possessing trait-level self-objectification (TSO) may interact to influence experiences of feeling like a body, or state self-objectification (SSO) [@fredrickson1998swimsuit, @garcia2016objectification]. Does this habit self-objectification lead women to also objectify other women, possibly to compare oneself?

@@

##Interpersonal Objectification
SIMO model and literature review [@gervais2020social]. Studies have shown that within social encounters women are gazed at more than men [@briton1995], often times feel ‘looked at’ within interpersonal interactions [@argyle1969; @miles2015stop], and will more than likely internalize the objectifying gaze on physical self [@young1979]. Moreover, perhaps the most adverse effect of objectifying treatment is that it effectively socializes girls and women to treat themselves as objects to be looked at and evaluated, an effect termed self-objectification  [@Bartky; @berger1972; @fredricksonetal1998]. We don't know about women-on-women interpersonal objectification. What might be different?

Self-objectification has been found to occur after a mere relationship prime among women [@sanchez2008romance] because in our culture women need to look attractive to obtain and maintain successful relationship, thus, the "college relationships condition may heighten self-objectification and the evaluation of other women in a sexualized way. Researchers have also studied interpersonal objectification and self-objectification in romantic relationships [@strelan2018birds]. Evidence is mixed, and women can sometimes have positive outcomes from appearance valuation within the context of an established romantic relationship and they experience less SSO after appearance comments within romantic relationships [@meltzer2020women], less is known about first dates. But this literature is also only about heterosexual relationships. A recent study has investigated the effect of self-objectification on the reduction of relationship building skills in general (including same-sex friendships) [CITE PWQ paper I reviewed, Yoder editor]. Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships [@brunelletal2010], adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance [@kahalon2018don; @quinnetal]. Both romantic relationships and relationships in general have been studies in the context of objectification, but neither of these literatures have investigated specifically same-gender relationships.

There is evidence that the experience of state self-objectification in mixed-gender contexts (stranger and romantic) and within the self in any moment (trying on a swimsuit) has negative consequences, but what about in the context of interaction with other women? Is the self-objectification experience in mixed-gender interactions the same as that experienced in same-gender interactions? The ample research demonstrating that the male gaze has a particularly detrimental effect would suggest no [@saguyetal2010, @calogero2004test; @gervais2011you; @gay2010my]. Women do objectify other women [@puvia2013being; @loughnan2015exploring; @harsey2020men], but when they do it, does it lead to self-objectification in the same way men's objectification of women does [@garcia2016objectification]? If so, does the self-objectification experienced in these same-gender interactions have the same negative consequences for authenticity in that ineraction?

##Interpersonal Objectification and Authenticity
There is some evidence that reduced authenticity is a consequence of self-objectification in the moment [@garcia2016objectification; again CITE PWQ paper I reviewed, Yoder editor]. This link has been justified by the literature on stigmatized-stigmatizer interactions, viewing the experience of being objectified in an interaction as an identity threat situation. When a woman is objectified by a man, and then experiences self-objectification, this intergroup encounter might trigger identitiy threat, but is the same negative consequence present when a woman is objectified by another woman? woman-woman interpersonal objectification processes might diverge from this process. One the one hand, she may not be having the same negative consequences the cascade from situations that threated group-based identity [@deaux1987putting; @dovidio2006nonverbal; @hebl2005promoting], but feeling like a body, rather than a full human, in an interaction may be enough to reduce feelings of authenticity and social competence, regardless of the gender of the objectifyer [@tolman2006looking; PWQ paper I reviewed, Yoder editor].
